### 2023-10-08 scripts/evaluation/experiment_factor.Rmd
update eval script

---
title: "Analysis of residual plot model"
output: html_document
date: "2023-10-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load_packages}
library(tensorflow)
library(keras)
library(tidyverse)
library(yardstick)
library(visage)
library(glue)
library(here)
```

```{r set_global_variables}
ALL_RES <- c(32L, 64L)
```

```{r load_model_and_test_set}
mod <- list()
test <- list()
for (res in ALL_RES) {
  mod[[res]] <- load_model_tf(here(glue("keras_tuner/best_models/experiment_factor/residual_plots/{res}")))
  test[[res]] <- flow_images_from_directory(here(glue("data/experiment_factor/residual_plots/{res}/mixed/test")),
                                            target_size = c(res, res),
                                            shuffle = FALSE)
}
```

```{r get_predictions}
pred <- list()
for (res in ALL_RES) {
  pred[[res]] <- mod[[res]]$predict(test[[res]]) %>%
    as.data.frame() %>%
    mutate(truth = test[[res]]$classes + 1) %>%
    mutate(truth = names(test[[res]]$class_indices)[truth]) %>%
    mutate(filenames = test[[res]]$filenames) %>%
    mutate(plot_uid = gsub(".*/(.*).png", "\\1", filenames)) %>%
    mutate(plot_uid = as.integer(plot_uid)) %>%
    mutate(pred = V1 > 0.5) %>%
    mutate(pred = ifelse(pred, "not_null", "null")) %>%
    mutate(pred = factor(pred)) %>%
    mutate(truth = factor(truth)) %>%
    mutate(res = res)
}
```

```{r get_meta}
meta <- readRDS(here("data/experiment_factor/residual_plots/meta.rds"))

effect_size_ref <- vi_survey %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  select(shape, e_sigma, a, b, n, x_dist, effect_size) %>%
  group_by(shape, e_sigma, a, b, n, x_dist) %>%
  summarise(effect_size = first(effect_size))
```

```{r merge_meta}
for (res in ALL_RES) {
  pred[[res]] <- pred[[res]] %>%
    left_join(meta) %>%
    left_join(effect_size_ref) %>%
    mutate(type = ifelse(!is.na(shape), "polynomial", "heteroskedasticity")) %>%
    mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete", x_dist))
}
```

# Basic summary

```{r basic_summary}

basic_summary <- function(dat, res = 32L) {
  dat %>%
  ggplot() +
  geom_bar(aes(pred, fill = type), position = "dodge") +
  geom_text(data = dat %>%
              group_by(truth, type, pred) %>%
              summarise(prop = n()/2000) %>%
              filter(type == "heteroskedasticity"),
  aes(pred, prop * 2000, label = format(prop, digits = 3)),
  nudge_x = -0.25, nudge_y = 30) +
  geom_text(data = dat %>%
              group_by(truth, type, pred) %>%
              summarise(prop = n()/2000) %>%
              filter(type == "polynomial"),
  aes(pred, prop * 2000, label = format(prop, digits = 3)),
  nudge_x = 0.25, nudge_y = 30) +
  xlab("Prediction") +
  theme_light() +
  facet_wrap(~truth, labeller = label_both) +
  ylab("Count") +
  scale_fill_brewer(palette = "Set2") +
  ggtitle("Summary of prediction", subtitle = glue("Image size {res} * {res}"))
}

for (res in ALL_RES) {
  print(basic_summary(pred[[res]], res))
}

```

```{r balance_accuracy}
pred %>%
  reduce(bind_rows) %>%
  group_by(res) %>%
  bal_accuracy(truth, pred) %>%
  kableExtra::kable(col.names = c("Input size", "Metric", "Estimator", "Estimate")) %>%
  kableExtra::kable_material()
```