### 2024-02-05 paper/paper.tex

% interactcadsample.tex
% v1.03 - April 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{lscape}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\usepackage{setspace}
\doublespacing

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\articletype{ARTICLE TEMPLATE}

\title{Automated assessment of residual plots with computer vision
models}

\author{\name{Weihao Li$^{a}$}
\affil{$^{a}$Department of Econometrics and Business Statistics, Monash
University, Clayton, VIC, Australia}
}

\thanks{CONTACT Weihao
Li. Email: \href{mailto:weihao.li@monash.edu}{\nolinkurl{weihao.li@monash.edu}}}

\maketitle

\begin{abstract}
TBD.
\end{abstract}

\begin{keywords}
TBD
\end{keywords}

\newpage
\tableofcontents
\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Plotting residuals is commonly regarded as a standard practice in linear
regression diagnostics
\citep[see][]{cook1982residuals, belsley1980regression}. This visual
assessment plays a crucial role in identifying deviations from model
assumptions, such as linearity, homoscedasticity, and normality. It also
helps in understanding the goodness of fit and various characteristics
of the model.

Generating a residual plot in most statistical software is often as
straightforward as executing a line of code or clicking a button.
However, accurately interpreting a residual plot can be challenging.
Consider Figure \ref{fig:false-finding} as an example, the residuals
display a triangular shape pointing to the left. While this might
suggest heteroskedasticity, it is important to avoid over-interpreting
this visual pattern. In this case, the fitted model is correctly
specified, and the triangular shape is actually a result of the skewed
distribution of the predictors, rather than indicating a flaw in the
model.

A residual plot can exhibit various visual features, but it is crucial
to recognize that some may arise from the characteristics of predictors
and the inherent randomness of the error, rather than indicating a
violation of model assumptions \citep{li2023plot}. The concept of visual
inference, as proposed by \citet{buja2009statistical}, provides an
inferential framework to assess whether residual plots indeed contain
visual patterns inconsistent with the model assumptions. The fundamental
idea involves testing whether the actual residual plot visually differs
significantly from null plots, which are created using residuals
generated from the null distribution. Typically, this is accomplished
through the lineup protocol. In this approach, the real residual plot is
embedded within a lineup alongside several null plots. If the real
residual plot can be distinguished from the lineup, it provides evidence
for rejecting the null hypothesis.

The practice of delivering a residual plot as a lineup is generally
regarded as a valuable approach. Beyond its application in residual
diagnostics, the lineup protocol has integrated into the analysis of
diverse subjects. For instance,
\cite{loy2013diagnostic, loy2014hlmdiag, loy2015you} illustrate its
applicability in diagnosing hierarchical linear models. Additionally,
\citet{widen2016graphical} demonstrates its utility in geographical
research, while \citet{krishnan2021hierarchical} explores its
effectiveness in forensic examinations.

However, as pointed out by \citet{li2023plot}, a primary limitation of
the lineup protocol lies in its reliance on human judgments. Unlike
conventional statistical tests that can be performed computationally in
statistical software, the lineup protocol requires human evaluation of
images. This characteristic makes it less suitable for large-scale
applications, given the associated high labor costs and time
requirements.

There is a substantial need to develop an approach that alleviates
analysts' workload by automating repetitive tasks and providing
standardized results in a controlled environment. The large-scale
evaluation of lineups is impractical without the use of technology and
machines.

The utilization of computers to interpret data plots has a rich history,
with early efforts such as ``Scagnostics'' by \citet{tukey1985computer},
focusing on scatterplot diagnostics. \citet{wilkinson2005graph} expanded
on this work, introducing graph theoretic scagnostics, which encompassed
nine computable measures applied to planar proximity graphs. These
measures, including, but not limited to, ``Outlying'', ``Skinny'',
``Stringy'', ``Straight'', ``Monotonic'', ``Skewed'', ``Clumpy'', and
``Striated'' aimed to characterize outliers, shape, density, trend,
coherence and other characteristics of the data. While this approach has
been inspiring, there is a recognition \citep{buja2009statistical} that
it may not capture all the necessary visual features that differentiate
actual residual plots from null plots. A more promising alternative
entails enabling computers to learn the function for extracting visual
features from residual plots. Essentially, this means empowering
computers to discern the crucial visual features for residual
diagnostics and determining the method to extract them. Modern computer
vision models are well-suited for addressing this challenge.

Modern computer vision models often rely on deep neural networks with
convolutional layers \citep{fukushima1982neocognitron}. These layers
leverage hierarchical patterns in data, downsizing and transforming
images by summarizing information in a small space. Numerous studies
have demonstrated the efficacy of convolutional layers in addressing
various vision tasks, including image recognition \citep{rawat2017deep}.
Despite the widespread use of computer vision models in fields like
computer-aided diagnosis \citep{lee2015image}, pedestrian detection
\citep{brunetti2018computer}, and facial recognition
\citep{emami2012facial}, their application in reading data plots remains
limited. While some studies have explored the use of computer vision
models for tasks such as reading recurrence plots for time series
regression \citep{ojeda2020multivariate}, time series classification
\citep{chu2019automatic, hailesilassie2019financial, hatami2018classification, zhang2020encoding},
anomaly detection \citep{chen2020convolutional}, and pairwise causality
analysis \citep{singh2017deep}, the application of reading residual
plots with computer vision models represents a relatively new field of
study.

In this paper, we develop computer vision models and integrate them into
the residual plots diagnostics workflow, filling the gap of\ldots. The
paper is structured as follows: \ldots{}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/false-finding-1}

}

\caption{An example residual vs fitted values plot (red line indicates 0). The vertical spread of the data points varies with the fitted values. This often indicates the existence of heteroskedasticity.}\label{fig:false-finding}
\end{figure}

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

\hypertarget{different-possible-configurations-of-the-model-formula}{%
\subsection{Different possible configurations of the model
formula}\label{different-possible-configurations-of-the-model-formula}}

There are various configurations of the computer vision model that can
be used to assess residual plots. We discuss these configurations below
based on two key components of the model formula: the input and the
output format.

\hypertarget{different-input-formats}{%
\subsubsection{Different input formats}\label{different-input-formats}}

Deep learning models are in general very sensitive to the input data.
The quality and relevance of the input data greatly influence the
model's capacity to generate insightful and meaningful results.

A straightforward approach involves feeding a vector of residuals along
with a vector of fitted values, essentially providing all the necessary
information for creating a residuals vs fitted values plot. However, a
drawback of this method is the dynamic input size, which changes based
on the number of observations. For modern computer vision models
implemented in mainstream software like TensorFlow
\citep{abadi2016tensorflow}, the input shape is typically fixed. One
solution is to pad the input vectors with leading or trailing zeros when
the input tensor expects longer vectors, but it may fail if the input
vector surpasses the designed length. Another strategy is to summarize
the residuals and fitted values separately using histograms and utilize
the counts as the input. By controlling the number of bins in the
histograms, it becomes possible to provide fixed-length input vectors.
Still, since the histogram only captures the marginal distribution of
residuals and fitted values respectively, it can not be used to
differentiate visual patterns with same marginal distributions but
different joint distributions.

Another approach is to use an image as input. The primary advantage of
using an image, as opposed to the vector format, is the utilization of
the sophisticated image processing architectures developed over the
years, such as the VGG16 architecture proposed in
\citet{simonyan2014very}. These architectures can effectively capture
and summarize spatial information from nearby pixels, which is less
straightforward with vector input. The only considerations are the image
resolution and the aesthetics of the residual plot. In general, higher
resolution provides more information to the model but comes with the
trade-off of increased complexity and greater difficulty in training. As
for the aesthetics of the residual plot, a practical solution is to
consistently present residual plots in the same style to the model. This
implies that the model can not accept arbitrary images as input but
requires the use of the same preprocessing software to convert provided
residuals and fitted values into a standardized style for the residual
plot.

Providing multiple residual plots to the model, such as a pair of plots,
a triplet or a lineup is also a possible option.
\citet{chopra2005learning} has shown that computer vision models
designed for image comparison can assess whether a pair of images are
similar or dissimilar. Applied to our specific problem, we can define
null plots of a fitted model to be similar to each other, while
considering actual residual plots to be distinct from null plots of any
fitted model. A triplet constitutes a set of three images, denoted as
\(image_1\), \(image_2\) and \(image_3\). It is often used to predict
whether \(image_2\) or \(image_3\) is more similar to \(image_1\),
proving particularly useful for establishing rankings between samples.
For this setup, we can apply the same criteria to define similarity
between images. However, it's important to note that these two
approaches usually require additional considerations regarding the loss
function and, at times, non-standard training processes due to shared
weights between different convolutional blocks.

Presenting a lineup to a model and asking it to predict which residual
plot is the most different one aligns closely with the lineup protocol.
However, if a lineup consists of a large amount of residual plots, the
resolution of the input image could become quite large, posing
challenges in training the model. This approach was experimented in a
pilot study conducted by us but the performance of the trained model was
sub-optimal.

We did not explore all the mentioned input formats due the considerable
costs associated with data preparation and model training. Considering
the implementation cost and the interpretability of the model, we
settled on the single residual plot input format.

\hypertarget{different-output-formats}{%
\subsubsection{Different output
formats}\label{different-output-formats}}

Given that the input is a single residual plot represented as a
fixed-resolution image, the output from the computer vision model can
take one of two formats: binary or numeric. This choice determines
whether the model belongs to a classification model or a regression
model. The binary outcome could be designed to indicate whether the
input image is a null plot, or whether the input image would be rejected
in a visual test conducted by humans. The latter option requires data
from prior human subject experiments, presenting difficulties in
controlling the quality of data due to variations in experimental
settings across different studies. Additionally, some visual inference
experiments are unrelated to linear regression models and residual plot
diagnostics, resulting in a limited amount of available training data.

Alternatively, the output could be designed as a meaningful numeric
measure, such as the average distance to good residual plots. This
approach necessitates defining a distance measure between residual
plots, which may be a non-trivial task. However, once a meaningful
distance measure is established, the prediction from the computer vision
model becomes an interpretable value with tangible significance. Studies
have demonstrated that defining a proper distance between images can
enhance the matching accuracy in image search compared to a binary
outcome model (ref here).

\hypertarget{distance-from-the-good-residual-plots}{%
\subsection{Distance from the good residual
plots}\label{distance-from-the-good-residual-plots}}

In a visual test, the observer will be asked to choose one or more plots
that stand out as most distinct from others in a given lineup. To
develop a computer vision model for assessing residual plots within the
visual inference framework, it is important to precisely define a
numerical measure of ``difference'' or ``distance'' between plots. This
distance can take the form of a basic statistical operation on pixels,
such as the sum of square differences. Alternatively, it could involve
established image similarity metrics like the Structural Similarity
Index Measure \citep{wang2004image}. The challenge lies in the fact that
metrics tailored for image comparison may not be suitable for evaluating
data plots, where only essential plot elements require assessment
\citep{chowdhury2018measuring}. Furthermore, scagnostics mentioned in
Section \ref{introduction} could be used to construct distance metrics
for data plots comparison, but the functional form still needs to be
carefully refined to accurately reflect the extent of the violations.

\hypertarget{residual-distribution}{%
\subsubsection{Residual distribution}\label{residual-distribution}}

The distance metrics proposed in this paper takes into account the fact
that we try to measure how different a residual plot is from a good
residual plot, or in other words, how different a given fitted model is
from a correctly specified model. For the classical normal linear
regression model, residuals \(\boldsymbol{e}\) are derived from the
fitted values \(\hat{\boldsymbol{y}}\) and observed values
\(\boldsymbol{y}\). Suppose the data generating process is known and the
regression model is correctly specified, by the Frisch-Waugh-Lowell
theorem \citep{frisch1933partial}, residuals \(\boldsymbol{e}\) can also
be treated as random variables and written as a linear transformation of
the error \(\boldsymbol{\varepsilon}\) formulated as
\(\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon}\), where
\(\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\)
is the residual operator and an idempotent matrix, \(\boldsymbol{X}\) is
the design matrix, \(\boldsymbol{I}_n\) is a \(n\) by \(n\) identity
matrix, and \(n\) is the number of observations.

One of the assumptions of the classical normal linear regression model
is the error \(\boldsymbol{\varepsilon}\) follows a multivariate normal
distribution with zero mean and constant variance, i.e.,
\(\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)\).
It can be known that residuals \(\boldsymbol{e}\) also follow a certain
probability distribution transformed from the multivariate normal
distribution, which will be denoted as \(Q\). This reference
distribution \(Q\) summarizes what good residuals should follow given
the predictors are known and fixed.

In ordinary least square, the minimization of the sum of square
residuals implies \(\sum_{i=1}^{n} e_i = 0\), making any residual value
to be a linear combination of the remaining \(n - 1\) residuals. This
effectively means \(\text{rank}(\boldsymbol{R}) = n - 1 < n\) and \(Q\)
is a degenerate multivariate distribution. To capture the
characteristics of \(Q\), such as moments, we can simulate a large
numbers of \(\boldsymbol{\varepsilon}\) and transform it to
\(\boldsymbol{e}\) to get the empirical estimates. For simplicity, we
replaced the variance-covariance matrix of residuals
\(\text{cov}(\boldsymbol{e}, \boldsymbol{e}) = \boldsymbol{R}\sigma^2\boldsymbol{R}' = \boldsymbol{R}\sigma^2\)
with a full-rank diagonal matrix
\(\text{diag}(\boldsymbol{R}\sigma^2)\), where \(\text{diag}(.)\) sets
the non-diagonal entries of a matrix to zeros. The resulting
distribution for \(\boldsymbol{Q}\) is
\(N(\boldsymbol{0}_n, \text{diag}(\boldsymbol{R}\sigma^2))\).

Distribution \(Q\) is derived from the correctly specified model.
However, if the model is misspecified, then the actual distribution of
residuals denoted as \(P\), will be different from \(Q\). For example,
if the data generating process contains variables correlated with any
column of \(\boldsymbol{X}\) but not included in \(\boldsymbol{X}\),
causing an omitted variable problem, \(P\) will be different from \(Q\)
because the residual operator obtained from the fitted model will not be
the same as \(\boldsymbol{R}\). Besides, if the
\(\boldsymbol{\varepsilon}\) follows a non-normal distribution such as a
multivariate lognormal distribution, \(P\) will usually be skewed and
has a long tail.

\hypertarget{kullback-leibler-divergence-of-p-from-q}{%
\subsubsection{\texorpdfstring{Kullback-Leibler divergence of \(P\) from
\(Q\)}{Kullback-Leibler divergence of P from Q}}\label{kullback-leibler-divergence-of-p-from-q}}

Define a proper distance between distributions is usually easier than
define a proper distance between data plots. Given the actual residual
distribution \(Q\) and the reference residual distribution \(P\), we
used a distance metric based on Kullback-Leibler divergence
\citep{kullback1951information} to quantify the difference between two
distributions

\begin{align}
\label{eq:kl-0}
D &= \log\left(1 + D_{KL}\right), \\
\label{eq:kl-1}
D_{KL} &= \int_{\mathbb{R}^{n}}\log\frac{p(\boldsymbol{e})}{q(\boldsymbol{e})}p(\boldsymbol{e})d\boldsymbol{e},
\end{align}

\noindent where \(p(.)\) is the probability density function for
distribution \(P\), and \(q(.)\) is the probability density function for
distribution \(Q\).

This distance metric was first proposed in \citet{li2023plot}. It was
mainly designed for measuring the effect size of non-linearity and
heteroskedasticity in a residual plot. \citet{li2023plot} showed that,
for a classical normal linear regression model that omits a necessary
higher-order predictors \(\boldsymbol{Z}\), and incorrectly assumes
\(\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)\)
while in fact
\(\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})\)
with \(\boldsymbol{V}\) being an arbitrary symmetric positive
semi-definite matrix, \(Q\) can be represented as
\(N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}))\).
Note that the variance-covariance matrix is replaced with the diagonal
matrix to ensure it is a full-rank matrix.

Since both \(P\) and \(Q\) are adjusted to be multivariate normal
distributions, equation \ref{eq:kl-1} can be further expanded to

\begin{align}
\label{eq:kl-2}
D_{KL} &= \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{W})|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{W})^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{W}))^{-1}\boldsymbol{\mu}_z\right),
\end{align}

\noindent where
\(\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z\),
and \(\boldsymbol{W} = \boldsymbol{R}\boldsymbol{V}\boldsymbol{R}\). The
assumed error variance \(\sigma^2\) is set to be
\(\text{tr}(\boldsymbol{V})/n\), which is the expectation of the
estimated variance.

\hypertarget{evaluation-of-kullback-leibler-divergence-for-non-normal-p}{%
\subsubsection{\texorpdfstring{Evaluation of Kullback-Leibler divergence
for non-normal
\(P\)}{Evaluation of Kullback-Leibler divergence for non-normal P}}\label{evaluation-of-kullback-leibler-divergence-for-non-normal-p}}

For non-normal error \(\boldsymbol{\varepsilon}\), the actual residual
distribution \(P\) is unlikely to be a multivariate normal distribution.
Thus, equation \ref{eq:kl-2} given in \citet{li2023plot} will not be
applicable to models with non-normality violations.

To evaluate the Kullback-Leibler divergence of non-normal \(P\) from
\(Q\), the fallback is to solve equation \ref{eq:kl-1} numerically.
However, since \(\boldsymbol{e}\) is a linear transformation of
non-normal random variables, it is very common that the general form of
\(P\) is unknown, meaning that we can not easily compute
\(p(\boldsymbol{e})\) using a well-known probability density function.
Additionally, even if \(p(\boldsymbol{e})\) can be calculated for any
\(\boldsymbol{e} \in \mathbb{R}^n\), it will be very difficult to do
numerical integration over the \(n\) dimensional space, because \(n\)
could be potentially very large.

In order to approximate \(D_{KL}\) in a practically computable manner,
the elements of \(\boldsymbol{e}\) are assumed to be independent of each
other. This assumption solves both of the issues mentioned above. First,
we no longer need to integrate over \(n\) random variables. The result
of equation \ref{eq:kl-1} is now the sum of the Kullback-Leibler
divergence evaluated for each individual residual thanks for the
independence assumption. Second, it is not required to know the joint
probability density \(p(\boldsymbol{e})\) any more. Instead, the
evaluation of Kullback-Leibler divergence for an individual residual
relies on the knowledge of the marginal density \(p_i(e_i)\), where
\(e_i\) is the \(i\)-th residual for \(i = 1, ..., n\). This is much
easier to approximate through simulation. It is also worth mentioning
that this independence assumption generally will not hold, since
\(\text{cov}(e_i, e_j) \neq 0\) if \(\boldsymbol{R}_{ij} \neq 0\) for
any \(1 \leq i < j \leq n\), but its existence is essential for reducing
the computational cost.

The algorithm for approximating equation \ref{eq:kl-1} starts from
simulating \(m\) sets of \(\boldsymbol{\varepsilon}\) according to the
error distribution. The simulated errors are stored in a matrix
\(\boldsymbol{A}\) with \(n\) rows and \(m\) columns, where each column
of \(\boldsymbol{A}\) is a set of realized values of
\(\boldsymbol{\varepsilon}\). Then, we can get \(m\) sets of realized
values of \(\boldsymbol{e}\) stored in the matrix \(\boldsymbol{B}\) by
applying the residual operator
\(\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}\). Furthermore, kernel
density estimation (KDE) with Gaussian kernel and optimal bandwidth
selected by the Silverman's rule of thumb \citep{silverman2018density}
is applied on each row of \(B\) to estimate \(p_i(e_i)\) for
\(i = 1, ..., n\). The KDE computation is done by the \texttt{density}
function in R.

Since the Kullback-Leibler divergence can be viewed as the expectation
of the log-likelihood ratio between distribution \(P\) and distribution
\(Q\) evaluated on distribution \(P\), we can reuse the simulated
residuals in matrix \(\boldsymbol{B}\) to estimate the expectation by
the sample mean. With the independence assumption, for non-normal \(P\),
\(D_{KL}\) can be approximated by

\begin{align}
\label{eq:kl-3}
D_{KL} &\approx \sum_{i = 1}^{n} \hat{D}_{KL}^{(i)}, \\
\hat{D}_{KL}^{(i)} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
\end{align}

\noindent where \(\hat{D}_{KL}^{(i)}\) is the estimator of the
Kullback-Leibler divergence for an individual residual \(e_i\),
\(B_{ij}\) is the \(i\)-th row and \(j\)-th column entry of the matrix
\(B\), \(\hat{p_i}(.)\) is the kernel density estimator of \(p_i(.)\),
\(q(.)\) is the normal density function with mean zero and an assumed
variance estimated as
\(\widehat{\sigma^2} = \sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)\),
and \(vec(.)\) is the vectorization operator which turns a
\(n \times m\) matrix into a \(nm \times 1\) column vector by stacking
the columns of the matrix on top of each other.

\hypertarget{approximation-of-the-distance-metric}{%
\subsubsection{Approximation of the distance
metric}\label{approximation-of-the-distance-metric}}

In the previous sections, we have defined a distance metric given in
equation \ref{eq:kl-0} for quantifying the difference between the actual
residual distribution \(P\) and an ideal reference distribution \(Q\).
You may have noticed that this distance metric can only be computed when
the data generating process is known. In reality, we often have no
knowledge about the data generating process, otherwise we do not need to
fit a linear regression model in the first place.

What we proposed in this paper is a method to approximate this distance
with a residual plot. Let \(D\) be the result of equation \ref{eq:kl-0}
indicating the extent of the model violations, and our estimator
\(\hat{D}\) is formulated as

\begin{equation}
\label{eq:d-approx}
\hat{D} = f_{CV}(V_{h \times w}(\boldsymbol{e}, \hat{\boldsymbol{y}})),
\end{equation}

\noindent where \(V_{h \times w}(.)\) is a plotting function that saves
a residuals vs fitted values plot with fixed aesthetic as an image with
\(h \times w\) pixels and three colour channels, \(f_{CV}(.)\) is a
computer vision model which takes an \(h \times w\) image as input and
predicts the distance in the domain \([0, +\infty)\).

With the approximated distance \(\hat{D}\), we will be able to know how
different the underlying distribution of the residuals is from a good
residual distribution. This is meaningful way to check if a residual
plot is a good residual plot, and to know the strength of the visual
signal embedded in the residual plot.

The approximated distance \(\hat{D}\) is not expected to be the same as
the original distance \(D\). This is largely because information
contained in a single residual plot is limited and it may not be able to
summarise all the characteristics of the residual distribution. For a
given residual distribution \(P\), we can generate many different
residual plots. Some of them share similar visual patterns, but some of
them could be visually very different from the rest, especially for
models with small \(n\). This suggests the error of the approximation
will vary depends on whether the observed residual plot is
representative or not.

\hypertarget{statistical-testing-based-on-the-approximated-distance}{%
\subsection{Statistical testing based on the approximated
distance}\label{statistical-testing-based-on-the-approximated-distance}}

\hypertarget{null-distribution-of-the-approximated-distance}{%
\subsubsection{Null distribution of the approximated
distance}\label{null-distribution-of-the-approximated-distance}}

Theoretically, the distance \(D\) for a correctly specified model is
\(0\), because \(P\) will be the same as \(Q\). However, the computer
vision model may not necessary predict \(0\) for a null plot. Using
Figure \ref{fig:false-finding} as an example, it contains a visual
pattern which is an indication of heteroskedasticity. We would not
expect the model to be able to magically tell if the suspicious pattern
is caused by the skewed distribution of the fitted values or the
existence of heteroskedasticity. Additionally, some null plots could
have outliers or strong visual patterns due to randomness, and a
reasonable model will try to summarise those information into the
prediction, resulting in \(\hat{D} > 0\).

This property is not an issue if \(\hat{D} \gg 0\) for which the visual
signal of the residual plot is very strong, and we usually do not need
any further examination of the significance of the result. However, if
the visual pattern is weak or moderate, having \(\hat{D}\) will not be
sufficient to determine if \(H_0\) that the model is correctly specified
should be rejected.

To solve this issue while aligning with the principle of visual
inference, the approximated distance \(\hat{D}\) can be viewed as a test
statistic. And the null distribution of this statistic can be
approximated by the empirical distribution of \(\hat{D}_{null}\) for a
large amount of null plots of a fitted model. The procedure involves
applying the residual rotation technique \citep{buja2009statistical} on
the fitted model to obtain null residuals. The null residuals are then
used to make null plots and fed into the model to get predictions. The
empirical distribution is constructed with the predicted distance.

There are two types of error associated with the approximation of the
null distribution. These include the sampling error and the estimation
error for model parameters. Increasing the number of null plots within
the distribution approximation can effectively reduce sampling error.
However, the estimation error for model parameters remains irreducible.
The residual rotation technique operates under the assumption of a
correct fitted model, using the estimated variance instead of the
unknown true variance to rescale rotated residuals. Consequently, under
\(H_0\), the true distribution of \(\hat{D}_{null}\) will slightly
deviate from the approximated null distribution.

\hypertarget{estimation-of-quantiles-of-the-null-distribution}{%
\subsubsection{Estimation of quantiles of the null
distribution}\label{estimation-of-quantiles-of-the-null-distribution}}

Let \(\hat{D}_{null}^{(i)}\) be the predicted distance of the \(i\)-th
null plots, where \(i = 1,...,n_{null}\) and
\(n_{null} \in \mathbb{N}^+\) is a sufficiently large number. Quantiles
of the null distribution can then be estimated using the sample
quantiles available in statistical software such as R. The details of
the sample quantile computation can be found in
\citet{hyndman1996sample}. In statistical testing, analysts often care
about certain quantiles of the null distribution, such as the 90\%
quantile, the 95\% quantile and the 99\% quantile. These quantiles are
used as thresholds to decide if \(H_0\) needs to be rejected. For
example, if \(\hat{D}\) is greater than and equal to the 95\% sample
quantile of the null distribution \(Q_{null}(0.95)\), we could say the
predicted distance for the actual residual plot is significantly
different from the predicted distance for null plots with 95\%
significance level. Based on our experience, in order to get a stable
estimate of the 95\% quantile, \(n_{null}\) usually needs to be at least
\(100\). And if the null distribution has a long tail, more null plots
will be needed. Alternatively, a p-value can be used to represents the
probability of observing an event equally or more extreme than the given
event under \(H_0\), and it can be estimated by
\(1/m\sum_{i=1}^{n_{null}}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)\).

\hypertarget{bootstrapping-the-approximated-distance}{%
\subsubsection{Bootstrapping the approximated
distance}\label{bootstrapping-the-approximated-distance}}

Bootstrap is often employed in linear regression when conducting
inference for estimated parameters. It is typically done by sampling
individual cases with replacement and refitting the regression model. If
the observed data accurately reflects the true distribution of the
population, the bootstrapped estimates can be used to measure the
variability of the parameter estimate without making strong
distributional assumptions about the data generating process.

Similarly, bootstrap can be applied on the approximated distance
\(\hat{D}\). For each refitted model \(M_{boot}^{(i)}\), there will be
an associated residual plot \(V_{boot}^{(i)}\) which can be fed into the
computer vision model to obtain \(\hat{D}_{boot}^{(i)}\), where
\(i = 1,...,n_{boot}\), and \(n_{boot}\) is the number of bootstrapped
samples. If we are interested in the variation of \(\hat{D}\), we can
use \(\hat{D}_{boot}^{(i)}\) to estimate a confidence interval.

Alternatively, since each \(M_{boot}^{(i)}\) has a set of estimated
coefficients \(\hat{\boldsymbol{\beta}}_{boot}^{(i)}\) and an estimated
variance \(\hat{\sigma^2}_{boot}^{(i)}\), a new approximated null
distribution can be construed and the corresponding 95\% sample quantile
\(Q_{boot}^{(i)}(0.95)\) can be computed. Then, if
\(\hat{D}_{boot}^{(i)} \geq Q_{boot}^{(i)}(0.95)\), \(H_0\) will be
rejected for \(M_{boot}^{(i)}\). The ratio of rejected
\(M_{boot}^{(i)}\) among all the refitted models provides an indication
of how often the assumed regression model are considered to be incorrect
if the data can be obtained repetitively from the same data generating
process. But this approach is computationally very expensive because it
requires \(n_{boot} \times n_{null}\) times of residual plot assessment.
In practice, \(Q_{null}(0.95)\) will be used to replace
\(Q_{boot}^{(i)}(0.95)\) in the computation.

\hypertarget{generation-of-training-data}{%
\subsection{Generation of training
data}\label{generation-of-training-data}}

\hypertarget{data-generating-process}{%
\subsubsection{Data generating process}\label{data-generating-process}}

While observational data is frequently employed in training models for
real-world applications, the data generating process of observational
data often remains unknown, making computation for our target variable
\(D\) unattainable. Consequently, the computer vision models developed
in this study were trained using synthetic data. This approach provides
us with precise label annotations. Additionally, it ensures a large and
diverse training dataset, as we have control over the data generating
process, and the simulation of the training data is relatively
cost-effective.

We have incorporated three types of residual departures of linear
regression model in the training data, including non-linearity,
heteroskedasticity and non-normality. All three departures can be
summarised by the data generating process formulated as

\begin{align}
\label{eq:data-sim}
\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 + \beta_2(\boldsymbol{z} + \beta_1\boldsymbol{w}) + \boldsymbol{k} \odot \boldsymbol{\varepsilon}, \\
\boldsymbol{z} &= \text{He}_j(g(\boldsymbol{x}_1, 2)), \\
\boldsymbol{w} &= \text{He}_j(g(\boldsymbol{x}_2, 2)), \\
\boldsymbol{k} &= \sqrt{\boldsymbol{1}_n + b(2 - |a|)(\boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 - a\boldsymbol{1}_n)^2},
\end{align}

\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}_1\),
\(\boldsymbol{x}_2\), \(\boldsymbol{z}\), \(\boldsymbol{w}\),
\(\boldsymbol{k}\) and \(\boldsymbol{\varepsilon}\) are vectors of size
\(n\), \(\boldsymbol{1}_n\) is a vector of ones of size \(n\),
\(\boldsymbol{x}_1\) and \(\boldsymbol{x}_2\) are two independent
predictors, \(\text{He}_j(.)\) is the \(j\)th-order probabilist's
Hermite polynomials \citep{hermite1864nouveau}, the \(\sqrt{(.)}\) and
\((.)^2\) operators are element-wise operators, \(\odot\) is the
Hadamard product, and \(g(., k)\) is a scaling function to enforce the
support of the random vector to be \([-k, k]^n\) defined as

\[g(\boldsymbol{x}, k) = 2k \cdot \frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k\boldsymbol{1}_n,~for~k > 0,\]
\noindent where \(x_{min} = \underset{i \in \{ 1,...,n\}}{min} x_i\),
\(x_{max} = \underset{i \in \{ 1,...,n\}}{max} x_i\) and \(x_i\) is the
\(i\)-th entry of \(\boldsymbol{x}\).

The residuals and fitted values of the fitted model is obtained by
regressing \(\boldsymbol{y}\) on \(\boldsymbol{x}_1\). This data
generating process is adopted from \citet{li2023plot} where it was used
to simulate residual plots with non-linearity and heteroskedasticity
visual patterns for human subject experiments.

In equation \ref{eq:data-sim}, predictor \(\boldsymbol{x}_1\) and
\(\boldsymbol{x}_2\) are independent of each other, so excluding
\(\boldsymbol{x}_2\) from the regression formula will not cause any
problem. However, \(\boldsymbol{z}\) and \(\boldsymbol{w}\) are
higher-order terms of \(\boldsymbol{x}_1\) and \(\boldsymbol{x}_2\). If
\(\beta_2 \neq 0\), the regression model will suffer from non-linearity
issues. Parameter \(j\) is a shape parameter controlling the number of
tuning points of the non-linearity pattern. Generally, greater values of
\(j\) will result in more tuning points.

Additionally, Scaling factor \(\boldsymbol{k}\) directly affects the
error distribution and it is correlated with \(\boldsymbol{x}_1\) and
\(\boldsymbol{x}_2\). If \(b \neq 0\) and
\(\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)\),
the constant variance assumption will be violated. Parameter \(a\) is a
shape parameter controlling the location of the smallest variance in a
residual plot.

Non-normality violations are introduced by defining a non-normal
distribution for \(\boldsymbol{\varepsilon}\).

\begin{itemize}
\tightlist
\item
  distribution of x1 and x2
\item
  summary table of all the factor values
\item
  example plots for the violations
\end{itemize}

\hypertarget{computation-of-scagnostics}{%
\subsubsection{Computation of
scagnostics}\label{computation-of-scagnostics}}

We have mentioned in Section \ref{introduction} that scagnostics are a
set of manually designed visual feature extraction functions. Although
our computer vision model will learn its own feature extraction function
during the training process, it will be beneficial to utilize additional
information provided by scagnostics to help computer vision model making
more accurate predictions.

For each generated residual plot, four scagnostics, namely,
``Monotonic'', ``Sparse'', ``Splines'' and ``Striped'' are computed
using the \texttt{cassowaryr} R package (ref here). The computed
measures along with the number of observations of the fitted model are
provided as the second input of the computer vision model. Other
scagnositcs are also informative, but unfortunately, they are
unavailable due to a fatal bug in the compiled C program of the
\texttt{interp} R package (ref here) that will crash the process in a
unpredictable manner. For reproducibility, we do not include these
scagnostics in the training data.

\hypertarget{crafting-a-balanced-training-set}{%
\subsubsection{Crafting a balanced training
set}\label{crafting-a-balanced-training-set}}

In order to train a robust computer vision model, we intentionally
control the distribution of the target variable \(D\) of the training
data. Making it uniform between \(0\) and \(7\). This is accomplished by
preparing \(50\) buckets each only accepts training samples with \(D\)
between \([7(i - 1)/49, 7i/49)\) for \(i < 50\), where \(i\) is the
index of the \(i\)-th bucket. For the \(50\)-th bucket, any training
samples with \(D >= 7\) will be accepted. We have prepared 80000
training images, so each bucket can only contain
\(80000 \div 50 = 1600\) training samples. The simulator will repeatedly
sample parameter values from the parameter sample, generate residuals
and fitted values using the data generating process, compute the
distance, and check if the sample can be accepted by the corresponding
bucket, until all the buckets are full.

\hypertarget{architecture-of-the-computer-vision-model}{%
\subsection{Architecture of the computer vision
model}\label{architecture-of-the-computer-vision-model}}

The architecture of the computer vision model is adopted from a existing
architecture called VGG16 that proven to be performant in image
classification \citep{simonyan2014very}. The first layer is the input
layer with shape \(n \times h \times w \times 3\) which could accept
\(n\) RGB images, followed by a grayscale conversion layer using the
luma formula under the CCIR 601 standard (ref here) to convert the
colour image to an grey scale image. Grayscale is enough for our task
since data points are plotted in black. Three different combinations of
\(h\) and \(w\) are chosen to seek a sufficient large image resolution
for this problem. This includes \(32 \times 32\), \(64 \times 64\) and
\(128 \times 128\).

The processed image is then used as the input of the first
convoluational block. The model has at most five consecutive
convolutional blocks which is similar to the original VGG16
architecture. In each block, there are two 2D convoluational layers
followed by two activation layers respectively, and there is a 2D max
pooling layer follows the second activation layer. The 2D convolutional
layer convolves the input with a fixed number of \(3 \times 3\)
convolution filters, and the 2D max pooling layer downsamples the input
along its spatial dimensions by taking the maximum value over a
\(2 \times 2\) window for each channel of the input. The activation
layer is set to be the rectified linear unit activation function (ReLu),
which is a standard practice in deep learning. This introduces a
non-linear transformation of the output of the 2D convolutional layer.
Furthermore, to regularize the training, a batch normalization layer is
added after each of the 2D convolutional layer and before the activation
layer. A dropout layer is also appended at the end of each convolutional
block to randomly set some inputs to zeros during training.

The output of the last convolutional block will be summarised by a
global max pooling layer or a global average pooling layer to gain a
two-dimensional tensor. Since we want to utilize the information
contained in scagnostics, the two-dimensional tensor is concatenated
with an additional \(n \times 5\) tensor, which contains the
``Monotonic'', ``Sparse'', ``Splines'' and ``Striped'' measures and the
number of observations for \(n\) residual plots.

The concatenated tensor will be fed into the final prediction block.
This block contains two fully-connected layers, the first one has at
least \(128\) units followed by a dropout layer. Sometimes, a batch
normalization layer will be inserted between the fully-connected layer
and the dropout layer. The second fully-connected layer has only one
unit, serves as the output of the model.

The model weights are randomly initialized and they are optimized by the
Adam optimizer with the mean square error loss function. Therefore, we
try to find a set of weights \(\hat{\boldsymbol{\theta}}\) by

\[\hat{\boldsymbol{\theta}} = \underset{\boldsymbol{\theta}}{argmin}\frac{1}{n_{train}}\sum_{i=1}^{n_{train}}(D_i - f_{\boldsymbol{\theta}}(V_i, S_i))^2,\]

\noindent where \(n_{train}\) is the number of training samples, \(V_i\)
is the \(i\)-th residual plot and \(S_i\) is the additional information
about the \(i\)-th residual plot including four scagnostics and the
number of observations.

\begin{itemize}
\tightlist
\item
  a diagram for the model
\end{itemize}

\hypertarget{training-and-hyperparameter-tuning}{%
\subsection{Training and hyperparameter
tuning}\label{training-and-hyperparameter-tuning}}

To obtain an optimal deep learning model, hyperparameters such as
learning rate often needs to be tuned via a tuner. In this study, we
used the Bayesian optimization tuner from the \texttt{keras-tuner}
Python library (ref here) to tune the hyperparameters. The complete list
of hyperparameters is provided in Table \ref{tab:hyperparameter}.

The number of base filters determines the number of filters for the
first 2D convolutional layer. In VGG16, the number of filters for the 2D
convolutional layer in a block is 2 times the number in the previous
block, except for the last block which has the same number of
convolution filters as the previous one. This hyperparameter helps to
control the complexity of the computer vision model. More base filters
will result in more trainable parameters.

Similarly, the number of units for the first fully-connected layer
determines the complexity of the final prediction block. And more units
will lead to more trainable parameters.

Dropout rate and batch normalization are flexible hyperparameters which
works jointly with other hyperparameters to smooth the training process.
A high dropout rate is needed when the model learns too much noise from
the training dataset. And a low dropout rate is needed when the model is
complex and difficult to converge. Batch normalization on the other hand
may be needed for solving the internal covariate shift problem due to
randomness in the weight initialization and randomness in the input
data.

Having additional inputs including the scagnostics and the number of
observations could be beneficial for making more accurate prediction. So
we let the tuner to decide if these inputs are actually needed.

Learning rate is one of the most important hyperparameters, and it
determines the step size of the optimization algorithm. A high learning
rate could help the model to avoid getting stuck in a local minimum, but
it could also ruin the optimization process by jumping over the global
minimum. A low learning rate could smooth the training process but it
increases the time to converge and the chance of getting stuck in local
minimum.

The model is trained on the M3 high-performance computing platform (ref
here) with TensorFlow (ref here) and Keras (ref here). During the
training process, 80\% of the training data are used as the actual
training data, and 20\% of the training data are reserved as validation
data. We let the Bayesian optimization tuner runs 100 trials to find the
best hyperparameter values according to the validation root mean square
error. The tuner will restore the best epoch of the best model from the
trials. Early stopping is also applied such that the training process
will halt if the validation root mean square error does not improve for
50 epochs. The maximum allowed epochs is 2000 where no models reach.

\begin{table}

\caption{\label{tab:hyperparameter}Name of hyperparameters and their correspoding domain for the computer vision model.}
\centering
\begin{tabular}[t]{ll}
\toprule
Hyperparameter & Domain\\
\midrule
Number of base filters & \{4, 8, 16, 32, 64\}\\
Dropout rate for convolutional blocks & {}[0.1, 0.6]\\
Batch normalization for convolutional blocks & \{false, true\}\\
Type of global pooling & \{max, average\}\\
Ignore additional inputs & \{false, true\}\\
\addlinespace
Number of units for the fully-connected layer & \{128, 256, 512, 1024, 2048\}\\
Batch normalization for the fully-connected layer & \{false, true\}\\
Dropout rate for the fully-connected layer & {}[0.1, 0.6]\\
Learning rate & {}[1e-8, 1e-1]\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{model-evaluation-methods}{%
\subsection{Model evaluation methods}\label{model-evaluation-methods}}

\begin{itemize}
\item
  RMSE for the test set
\item
  R\^{}2
\item
  Mean bias deviation to understand the overall bias
\item
  quantile loss to understand how well the model captures the entire
  distribution of D
\item
  Percentage of predictions within a tolerance interval (like 0.1)
\item
  comparison with human visual inference
\item
  overview of the human subject experiment
\item
  metrics
\item
  agreement rate
\item
  check cases that are aggreed
\item
  check cases that are disaggreed
\end{itemize}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{optimized-hyperparameters}{%
\subsection{Optimized hyperparameters}\label{optimized-hyperparameters}}

Based on the tuning process described in Section
\ref{training-and-hyperparameter-tuning}, the optimized hyperparameter
values are provided in Table \ref{tab:best-hyperparameter}. It can be
observed that at least \(32\) base filters are needed, and the
preferable choice for the \(64 \times 64\) model and the
\(128 \times 128\) model is \(64\) base filters, which is the same as
the original VGG16 architecture. The optimized values for dropout rate
for convolutional blocks is around \(0.4\). And applying batch
normalization for convolutional blocks has a positive impact on the
performance. All optimized models choose to preserve the additional
inputs to reduce the validation error. The number of units needed for
the fully-connected layer is \(256\), which is not a very large number
comparing to the VGG16 classifier, suggesting that the problem we are
trying to solve is relatively easy. The optimized learning rates are
usually smaller than the default value recommended by Keras which is
0.001.

\begin{table}

\caption{\label{tab:best-hyperparameter}Hyperparameters values for the optimized computer vision models with different input sizes.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llll}
\toprule
Hyperparameter & $32 \times 32$ & $64 \times 64$ & $128 \times 128$\\
\midrule
Number of base filters & 32 & 64 & 64\\
Dropout rate for convolutional blocks & 0.4 & 0.3 & 0.4\\
Batch normalization for convolutional blocks & true & true & true\\
Type of global pooling & max & average & average\\
Ignore additional inputs & false & false & false\\
\addlinespace
Number of units for the fully-connected layer & 256 & 256 & 256\\
Batch normalization for the fully-connected layer & false & true & true\\
Dropout rate for the fully-connected layer & 0.2 & 0.4 & 0.1\\
Learning rate & 0.0003 & 0.0006 & 0.0052\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{model-performance}{%
\subsection{Model performance}\label{model-performance}}

The training and test performance of optimized models with three
different input sizes are provided in Table \ref{tab:performance}. Among
these models, the \(64 \times 64\) model and the \(32 \times 32\) model
are consistently having the best training and test performance
respectively. The mean absolute error shows that the difference between
\(\hat{D}\) and \(D\) is around \(0.43\) on the test set, which is very
small considering the range of \(D\) is normally between \(0\) and
\(7\). The high \(R^2\) also suggests the prediction is mostly linearly
correlated with the target.

Figure \ref{fig:model-performance} displays a hexagonal heatmap for
\(D - \hat{D}\) vs \(D\). For residual plots with \(D\) close to zero,
all the optimized models tends to over-predict \(\hat{D}\), leading to
negative residuals. When \(D\) is between \(1.5\) and \(6\), the smooth
curves indicate that \(\hat{D}\) is not heavily biased, and there are
some instances where \(\hat{D}\) is severely under-predicted. And when
\(D\) is greater than \(6\), there are more instances being
under-predicted than over-predicted.

Given the model performance metrics, only the best model evaluated on
the test set, i.e.~the \(32 \times 32\) model, will be used in the
following analysis.

\begin{table}

\caption{\label{tab:performance}The training and test performance of three optimized models with different input sizes.}
\centering
\begin{tabular}[t]{lllll}
\toprule
 & RMSE & $R^2$ & MAE & Huber loss\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Training set}}\\
\hspace{1em}$32 \times 32$ & 0.531 & 0.937 & 0.364 & 0.126\\
\hspace{1em}$64 \times 64$ & 0.405 & 0.963 & 0.260 & 0.072\\
\hspace{1em}$128 \times 128$ & 0.432 & 0.959 & 0.290 & 0.084\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Test set}}\\
\hspace{1em}$32 \times 32$ & 0.660 & 0.901 & 0.434 & 0.181\\
\hspace{1em}$64 \times 64$ & 0.674 & 0.897 & 0.438 & 0.186\\
\hspace{1em}$128 \times 128$ & 0.692 & 0.892 & 0.460 & 0.199\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/model-performance-1}

}

\caption{Hexagonal heatmap for residuals vs distance on training and test data for three optimized models with different input sizes. The blue lines are smoothing curves produced by fitting gnealized additive models.}\label{fig:model-performance}
\end{figure}

\hypertarget{comparison-with-human-visual-inference}{%
\subsection{Comparison with human visual
inference}\label{comparison-with-human-visual-inference}}

\hypertarget{overview-of-the-human-subject-experiment}{%
\subsubsection{Overview of the human subject
experiment}\label{overview-of-the-human-subject-experiment}}

In order to check the validity of the proposed computer vision model,
residual plots presented in the human subject experiment conducted by
\citet{li2023plot} will be assessed. This study has collected \(7974\)
human responses to \(1152\) lineups. Each lineup contains one randomly
placed actual residual plot and 19 null plots. Among the \(1152\)
lineups, \(24\) are attention check lineups in which the visual patterns
are designed to be extremely obvious and very different from the
corresponding to null plots, \(36\) are null lineups where all the
lineups consist of only null plots, \(279\) are lineups with uniform
predictor distribution evaluated by \(11\) participants, and the
remaining \(813\) are lineups with discrete, skewed or normal predictor
distribution evaluated by \(5\) participants. Attention check lineups
and null lineups will not be assessed in the following analysis.

In \citet{li2023plot}, the residual plots are simulated from a data
generating process which is a special case of Equation
\ref{eq:data-sim}. The main characteristic is the model violations are
introduced separately, meaning non-linearity and heteroskedasticity will
not co-exist in one lineup but assigned uniformly to all lineups.
Additionally, non-normality and multiple predictors are not considered
in their experimental design.

\hypertarget{model-performance-on-the-human-data}{%
\subsubsection{Model performance on the human
data}\label{model-performance-on-the-human-data}}

For each lineup used in \citet{li2023plot}, there are one actual
residual plot and \(19\) null plots. While the distance \(D\) for the
actual residual plot depends on the underlying data generating process,
the distance \(D\) for the null plots are zeros. We have used our
optimized computer vision model to predict \(\hat{D}\) for both the
actual residual plots and the null plots.

The performance metrics of \(\hat{D}\) for actual residual plots are
provided in Table \ref{tab:experiment-performance}. It can be observed
that all the performance metrics are slightly worse than the metrics
evaluated on the test data. Still, the mean absolute error remains at a
low level, and the linearly correlation between the prediction and the
true value is still very high. Lineups with non-linearity issues are
more difficult to be predicted than those with heteroskedasticity
issues.

Figure \ref{fig:rd-human} displays the \(D - \hat{D}\) vs \(D\) plot for
actual residual plots used in the experiment. The behaviour of the
prediction is roughly the same as what we can be observed from the
evaluation of the test data. The model over-predicts \(\hat{D}\) when
\(D\) is close to zero, and under-predicts \(\hat{D}\) when \(D\) is
greater than \(6\). However, for residual plots with non-linearity
issues, we do not observe the over-prediction for \(D > 6\) because
there are not many cases in that area. And the over-prediction happens
at around \(2 < D < 4\).

For null plots, a histogram of the prediction is provided in Figure
\ref{fig:hist-null-human}. In most cases, \(\hat{D}\) will be predicted
at around \(1\). And as \(\hat{D}\) increases, the number of predictions
decreases. There are some outliers with \(\hat{D} > 5\).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/rd-human-1}

}

\caption{Residuals vs distance plot on lineups used in the human subject experiment. The blue line is a smoothing curve produced by fitting a gnealized additive model.}\label{fig:rd-human}
\end{figure}

\begin{table}

\caption{\label{tab:experiment-performance}The performance of the $32 \times 32$ model on the data used in the human subject experiment.}
\centering
\begin{tabular}[t]{lllll}
\toprule
Voilation & RMSE & $R^2$ & MAE & Huber loss\\
\midrule
heteroskedasticity & 0.721 & 0.852 & 0.553 & 0.235\\
non-linearity & 0.738 & 0.770 & 0.566 & 0.246\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/hist-null-human-1}

}

\caption{Histogram of predicted distance on null plots used in the human subject experiment.}\label{fig:hist-null-human}
\end{figure}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-3-1} \end{center}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-4-1}

}

\caption{A weighted detection rate vs $\delta$-differnence plot.}\label{fig:unnamed-chunk-4}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-5-1}

}

\caption{A weighted detection rate vs $\gamma$-number of extreme nulls plot.}\label{fig:unnamed-chunk-5}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-6-1}

}

\caption{A weighted detection rate vs $\delta$-differnence plot for lineups evaluated by 11 people.}\label{fig:unnamed-chunk-6}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-7-1}

}

\caption{A weighted detection rate vs $\gamma$-number of extreme nulls plot for lineups evaluated by 11 people.}\label{fig:unnamed-chunk-7}
\end{figure}

\hypertarget{when-the-model-works}{%
\subsection{When the model works}\label{when-the-model-works}}

\begin{itemize}
\tightlist
\item
  simple examples (non-linearity, heteroskedasticity, \ldots)
\item
  datasaurus
\end{itemize}

\hypertarget{when-the-model-does-not-work}{%
\subsection{When the model does not
work}\label{when-the-model-does-not-work}}

\begin{itemize}
\tightlist
\item
  human detect but model does not
\item
  cartoon residuals?
\end{itemize}

\hypertarget{workflow-how-one-use-this-model-small-showcase}{%
\subsection{Workflow: how one use this model? (small
showcase)}\label{workflow-how-one-use-this-model-small-showcase}}

\hypertarget{dicussion}{%
\section{Dicussion}\label{dicussion}}

There are other kinds of residual departures like autocorrelation that
are not considered in this study. The primary goal of this paper is to
establish a new way of evaluating residual plot and conducting visual
test with computer vision models. Building computer vision models for
other model violations and other types of diagnostic plots could be
future directions of this field.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\begin{itemize}
\tightlist
\item
  Summary of findings
\item
  Contributions to the field
\item
  Future directions for research
\end{itemize}

\bibliographystyle{tfcad}
\bibliography{ref.bib}

\end{document}