### 2024-05-13 paper/appendix.tex

% interactcadsample.tex
% v1.03 - April 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{lscape}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\usepackage{setspace}
\doublespacing

\begin{document}

\articletype{Draft paper (Incomplete)}

\title{Appendix: Automated assessment of residual plots with computer
vision models}

\author{\name{Weihao Li$^{a}$, Dianne Cook$^{a}$, Emi Tanaka$^{b,
c}$, Susan VanderPlas$^{d}$, Klaus Ackermann$^{a}$}
\affil{$^{a}$Department of Econometrics and Business Statistics, Monash
University, Clayton, VIC, Australia; $^{b}$Biological Data Science
Institute, Australian National University, Acton, ACT,
Australia; $^{c}$Research School of Finance, Actuarial Studies and
Statistics, Australian National University, Acton, ACT,
Australia; $^{d}$Department of Statistics, University of Nebraska,
Lincoln, Nebraska, USA}
}

\thanks{CONTACT Weihao
Li. Email: \href{mailto:weihao.li@monash.edu}{\nolinkurl{weihao.li@monash.edu}}, Dianne
Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}, Emi
Tanaka. Email: \href{mailto:emi.tanaka@anu.edu.au}{\nolinkurl{emi.tanaka@anu.edu.au}}, Susan
VanderPlas. Email: \href{mailto:susan.vanderplas@unl.edu}{\nolinkurl{susan.vanderplas@unl.edu}}, Klaus
Ackermann. Email: \href{mailto:Klaus.Ackermann@monash.edu}{\nolinkurl{Klaus.Ackermann@monash.edu}}}

\maketitle

\begin{abstract}
TBD.
\end{abstract}

\begin{keywords}
TBD
\end{keywords}

\section{Additional details about data
generation}\label{additional-details-about-data-generation}

\subsection{Computation of
scagnostics}\label{computation-of-scagnostics}

In Section \ref{introduction}, we mentioned that scagnostics consist of
a set of manually designed visual feature extraction functions. While
our computer vision model will learn its own feature extraction function
during training, leveraging additional information from scagnostics can
enhance the model's predictive accuracy.

For each generated residual plot, we computed four scagnostics --
``Monotonic,'' ``Sparse,'' ``Splines,'' and ``Striped'' -- using the
\texttt{cassowaryr} R package \citep{mason2022cassowaryr}. These
computed measures, along with the number of observations from the fitted
model, were provided as the second input for the computer vision model.
Although other scagnostics are informative, they are currently
unavailable due to a fatal bug in the compiled C program of the
\texttt{interp} R package \citep{Albrecht2023interp}, which may
unpredictably crash the process. For reproducibility, we excluded these
scagnostics from the training data.

\section{Neural Network Layers Used in the
Study}\label{neural-network-layers-used-in-the-study}

This study employs seven types of neural network layers, all of which
are standard components frequently found in modern deep learning models.
These layers are well-documented in textbooks like
\citet{goodfellow2016deep}, which offer thorough explanations and
mathematical insights. In this section, we will offer a concise overview
of these layers, drawing primarily from the insights provided in
\citet{goodfellow2016deep}.

\subsection{Dense Layer}\label{dense-layer}

The Dense layer, also known as the fully-connected layer, is the
fundamental unit in neural networks. It conducts a matrix multiplication
operation between the input matrix \(\boldsymbol{X}\) and a weight
matrix \(\boldsymbol{W}\) to generate the output matrix
\(\boldsymbol{O}\), which can be written as

\[\boldsymbol{O} = \boldsymbol{X}\boldsymbol{W} + b,\]

where \(b\) is the intercept.

\subsection{ReLu Layer}\label{relu-layer}

The ReLU layer, short for rectified linear unit, is an element-wise
non-linear function introduced by \citet{nair2010rectified}. It sets the
output elements to zero if the corresponding input element is negative;
otherwise, it retains the original input. Mathematically, it can be
expressed as:

\[\boldsymbol{O}(i,j) = max\{0, \boldsymbol{X}(i,j)\},\]

where \(\boldsymbol{O}(i,j)\) is the \(i\)th row and \(j\)th column
entry of matrix \(\boldsymbol{O}\), and \(\boldsymbol{X}(i,j)\) is the
\(i\)th row and \(j\)th column entry of matrix \(\boldsymbol{X}\).

\subsection{Convolutaional Layer}\label{convolutaional-layer}

In Dense layers, matrix multiplication leads to each output unit
interacting with every input unit, whereas convolutional layers operate
differently with sparse interactions. Here, an output unit in a
convolutional layer is connected solely to a subset of input units, and
the weight is shared across all input units. Achieving this involves
using a kernel, typically a small square matrix, to conduct matrix
multiplication across all input units. Precisely, this concept can be
formulated as:

\[O(i, j) = \sum_m\sum_nI(i - m, j - n)K(m, n),\]

where \(m\) and \(n\) are the row and columns indices of the kernel
\(K\).

If there are multiple kernels used in one covolutional layer, then each
kernel will have its own weights and the output will be a
three-dimensional tensor, where the length of the third channel is the
number of kernels.

\subsection{Pooling Layer}\label{pooling-layer}

A pooling layer substitutes the input at a specific location with a
summary statistic derived from nearby input units. Typically, there are
two types of pooling layers: max pooling and average pooling. Max
pooling computes the maximum value within a rectangular neighborhood,
while average pooling calculates their average. Pooling layers helps
making the representation approximately invariant to minor translations
of the input. The output matrix of a pooling layer is approximately
\(s\) times smaller than the input matrix, where \(s\) represents the
length of the rectangular area. This can be formulated as:

\[O(i, j) = \underset{m,n}{\max} I(si + m,sj+n).\]

\subsection{Global Pooling Layer}\label{global-pooling-layer}

A global pooling layer condenses an input matrix into a scalar value by
either extracting the maximum or computing the average across all
elements. This layer acts as a crucial link between the convolutional
structure and the subsequent dense layers in a neural network
architecture. When convolutional layers utilize multiple kernels, the
output becomes a three-dimensional tensor with numerous channels. In
this scenario, the global pooling layer treats each channel
individually, much like distinct features in a conventional classifier.
This approach facilitates the extraction of essential features from
complex data representations, enhancing the network's ability to learn
meaningful patterns.

\subsection{Batch Normalization Layer}\label{batch-normalization-layer}

\subsection{Dropout Layer}\label{dropout-layer}

\bibliographystyle{tfcad}
\bibliography{ref.bib}

\end{document}